

\chapter*{Acknowledgements}
\thispagestyle{empty}

Thank you so much for choosing this text and a big thanks to all who made this possible!

Kenneth would like to thank Front Range Community College for their relentless dedication to excellence, inclusivity, and student success. To help Ken help our students, they supported this project with an Innovation Grant. In particular, thanks to FRCC president Andy Dorsey for believing in the project and Christy Gomez for her vision, structure, and pragmatism.  He would like to thank his family Ken, Gina, Maria, Bryan, and Keenan for the (even in a family that loves combinatorics) countless mathematical adventures throughout the years.  He is also grateful for the training he received from generous mentors while he was just a lowly graduate student at Colorado State University -- especially from Alexander Hulpke and Tim Penttila.  Lastly, thanks to his wife Faith for all the love and support throughout the project. She learned to make coffee just so she could help, despite thinking it's the most disgusting liquid in the visible universe.  

Jenna would like to thank the fantastic faculty at Front Range Community College -- Boulder County Campus and University of Colorado Boulder for their training and dedication to helping her become a better mathematician.
She would also like to thank the Monks family who has done so much to help her mathematically and encouraged her to continue to grow.
Finally, thanks to her family for brewing the coffee and putting up with her hermit tendencies.

%Jenna would also like to thank our lord and savior of darkness.. Cthulhu. May your tentacles always find proper destruction.
%lol cthulhu is such a great word. the spelling and the phonetics.
%I know right! Plus he's a city eater with tentacles. How is that not cool?

Thanks to Joe Darschewski, FRCC-BCC's excellent Calculus instructor, for his being the first instructor to use the text and offer an enormous amount of helpful feedback.  He also typeset the new Mixed Practice sections, which will be very helpful to the students!

We hope you have fun with this resource and find it helpful!  It is a beautiful subject, and we tried to honor that with a beautiful text.  The text is still in its infancy, and we welcome any and all feedback you could give us.  Thank you in advance for any comments, complaints, suggestions, and questions. \\

\noindent Your book makers, \\
	Kenneth M Monks, Front Range Community College - Boulder County Campus 
	\begin{verbatim}
    kenneth.monks@frontrange.edu
    \end{verbatim}
	Jenna M Allen, University of Colorado Boulder 
	\begin{verbatim}
    jenna.m.allen@colorado.edu
    \end{verbatim}
%Why would email be allowed at a renaissance fair? lol And yes they will eat you if you are a meanie pants. There is a reason you'll never hear about my ex boyfriends... teehee.

%LOL!  Nice.  At least they had one good purpose.
%Finest dragon fodder you can find.
\clearpage